---
title: "DC Property ML — Exploratory Data Analysis"
author: "Aaron Niecestro"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 150)
library(here)
source(here("src",   "01_load_config.R"))
source(here("src",   "02_load_data.R"))
source(here("src",   "03_split_data.R"))
source(here("utils", "utils_preprocessing.R"))
source(here("utils", "utils_plots.R"))

cfg   <- load_config()
set_project_seed(cfg)
```

## 1. Load & Clean Data

```{r load}
df <- load_raw_data(cfg, verbose = TRUE)
dim(df)
```

## 2. Data Quality Report

```{r quality}
dq <- data_quality_report(df)
knitr::kable(dq, caption = "Data Quality Summary")
```

## 3. Descriptive Statistics

These match the summary statistics reported in the paper (Section: Data Description).

```{r desc-stats}
num_summ <- numeric_summary(df)
knitr::kable(num_summ, digits = 4,
             caption = "Descriptive Statistics — Numeric Predictors")
```

Key values from the paper:

| Variable    | Paper Mean | Paper Range         |
|-------------|-----------|---------------------|
| Price_10K   | 57.725    | 1.027 – 910.000     |
| Bathrooms   | 2.522     | 0 – 11              |
| Rooms       | 7.438     | 0 – 30              |
| Bedrooms    | 3.422     | 0 – 15              |
| AYB_age     | 84.58     | 1 – 254             |
| EYB_age     | 49.24     | 1 – 86              |
| Remodel_age | 8.398     | 0 – 139             |

## 4. Price Distribution

```{r price-hist, fig.cap="Dashed red line = mean (57.725), dashed green = median (44.365)"}
save_price_histogram(df, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_price_histogram.png"))
```

The distribution is right-skewed, consistent with the paper's note that most properties
cluster below the mean. The two classification thresholds (mean and median) are overlaid.

## 5. Price by Categorical Variables

```{r price-by-grade}
save_price_by_category(df, "Grade", cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_price_by_grade.png"))
```

```{r price-by-condition}
save_price_by_category(df, "Condition", cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_price_by_condition.png"))
```

```{r price-by-ward}
save_price_by_category(df, "Ward", cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_price_by_ward.png"))
```

## 6. Geographic Distribution

```{r geo-map}
save_geo_price_map(df, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_geo_price_map.png"))
```

## 7. Correlation Heatmap

```{r corr-heatmap}
save_correlation_heatmap(df, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "eda_correlation_heatmap.png"))
```

## 8. Train / Test Split

```{r split}
splits  <- make_split(df, cfg)
train   <- splits$train
test    <- splits$test

cat(sprintf("Train: %d rows | Test: %d rows\n", nrow(train), nrow(test)))
cat(sprintf("Train mean Price_10K: %.4f\n", mean(train$Price_10K)))
cat(sprintf("Test  mean Price_10K: %.4f\n", mean(test$Price_10K)))
```

```{r class-labels}
source(here("src", "03_split_data.R"))
labeled <- add_classification_labels(train, test, cfg)
cat("\nClass balance (mean split — train):\n")
print(table(labeled$train$class_mean))
cat("\nClass balance (median split — train):\n")
print(table(labeled$train$class_median))
```
