---
title: "DC Property ML — Regression & Classification Models"
author: "Aaron Niecestro"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 150)
library(here)

# Source all modules
source(here("src",   "01_load_config.R"))
source(here("src",   "02_load_data.R"))
source(here("src",   "03_split_data.R"))
source(here("src",   "04_feature_engineering.R"))
source(here("src",   "05_models_linear.R"))
source(here("src",   "06_models_polynomial.R"))
source(here("src",   "07_models_knn.R"))
source(here("src",   "08_models_lda_qda.R"))
source(here("src",   "09_models_ridge_lasso.R"))
source(here("src",   "10_models_pcr_pls.R"))
source(here("src",   "11_models_trees.R"))
source(here("src",   "12_models_random_forest.R"))
source(here("src",   "13_models_svm.R"))
source(here("src",   "14_evaluate.R"))
source(here("utils", "utils_io.R"))
source(here("utils", "utils_plots.R"))
source(here("utils", "utils_preprocessing.R"))

cfg <- load_config()
set_project_seed(cfg)
ensure_output_dirs(cfg)
```

## 1. Load Splits

```{r load-splits}
splits  <- load_splits(cfg)
train   <- splits$train
test    <- splits$test
labeled <- add_classification_labels(train, test, cfg)
train_c <- labeled$train
test_c  <- labeled$test

cat(sprintf("Train: %d rows | Test: %d rows\n", nrow(train), nrow(test)))
```

## 2. Feature Engineering

```{r features}
train_poly <- add_polynomial_features(train, cfg)
test_poly  <- add_polynomial_features(test,  cfg)
train_Xy   <- build_Xy(train_poly)
test_Xy    <- build_Xy(test_poly)

cat(sprintf("Polynomial feature matrix: %d columns\n", ncol(train_Xy$X)))
```

---

## 3. Regression Models

### 3.1 Simple & Multiple Linear Regression

```{r linear}
lm_model <- fit_linear(train)
summary(lm_model)
```

**Variable selection (AIC / BIC):**

```{r variable-selection, eval=FALSE}
# These are slow on the full dataset — run interactively
lm_aic <- select_by_aic(train)
lm_bic <- select_by_bic(train)
```

```{r linear-diag}
save_residual_plots(lm_model, cfg, tag = "linear_diagnostics")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "linear_diagnostics.png"))
```

```{r vif}
vif_vals <- check_vif(lm_model)
knitr::kable(data.frame(Predictor = names(vif_vals), VIF = round(vif_vals, 3)),
             caption = "Variance Inflation Factors — Linear Model")
```

### 3.2 Polynomial Regression

Polynomial powers from paper: Bathrooms^8, Bedrooms^8, Fireplaces^4,
AYB_age^5, EYB_age^4, Remodel_age^4 + Bathrooms × Condition interactions.

```{r polynomial, cache=TRUE}
poly_model  <- fit_polynomial(train, cfg)
summary(poly_model)
```

```{r poly-outliers}
outlier_idx   <- detect_outliers(poly_model)
cat(sprintf("Bonferroni outliers found: %d\n", length(outlier_idx)))

poly_refitted <- refit_without_outliers(train, cfg, outlier_idx)
poly_model    <- poly_refitted$model
```

```{r poly-diag}
save_diagnostic_plots(poly_model, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "polynomial_diagnostics.png"))
```

### 3.3 Ridge Regression

```{r ridge, cache=TRUE}
ridge_model <- fit_ridge(train_Xy, cfg)
save_cv_plot(ridge_model, cfg, "ridge_cv")
save_coef_path_plot(ridge_model, cfg, "ridge_coef_path")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "ridge_cv_lambda.png"))
```

```{r ridge-eval}
ridge_eval <- eval_ridge(ridge_model, test_Xy)
cat(sprintf("Ridge Test MSE: %.2f\n", ridge_eval$mse))
```

### 3.4 Lasso Regression

```{r lasso, cache=TRUE}
lasso_model <- fit_lasso(train_Xy, cfg)
save_cv_plot(lasso_model, cfg, "lasso_cv")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "lasso_cv_lambda.png"))
```

```{r lasso-eval}
lasso_eval <- eval_lasso(lasso_model, test_Xy)
cat(sprintf("Lasso Test MSE: %.2f | Non-zero coefs: %d\n",
            lasso_eval$mse, lasso_eval$n_nonzero))
compare_ridge_lasso(ridge_eval, lasso_eval)
```

### 3.5 Principal Component Regression (PCR)

```{r pcr, cache=TRUE}
pcr_model <- fit_pcr(train_poly, cfg)
save_rmsep_plot(pcr_model, cfg, "pcr_rmsep")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "pcr_rmsep.png"))
pcr_eval  <- eval_pcr(pcr_model, test_poly)
```

### 3.6 Partial Least Squares (PLS)

```{r pls, cache=TRUE}
pls_model <- fit_pls(train_poly, cfg)
save_rmsep_plot(pls_model, cfg, "pls_rmsep")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "pls_rmsep.png"))
pls_eval  <- eval_pls(pls_model, test_poly)
```

*Paper finding: RMSEP stabilises at ~3 components for PLS vs. 29 for PCR.*

### 3.7 Regression Tree + Pruning

```{r tree}
tree_model  <- fit_tree(train, cfg)
save_cp_plot(tree_model, cfg)
pruned_tree <- prune_tree(tree_model, cfg)
save_tree_plot(pruned_tree, cfg, "tree_pruned")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "tree_pruned.png"))
tree_eval   <- eval_tree(pruned_tree, test)
cat(sprintf("Pruned Tree Test MSE: %.4f\n", tree_eval$mse))
```

### 3.8 Random Forest & Bagging

```{r rf, cache=TRUE}
rf_model   <- fit_random_forest(train, cfg, ntree = 500)
save_rf_error_plot(rf_model, cfg, "rf_oob_error_full")
save_importance_plot(rf_model, cfg, "rf_importance_full")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "rf_importance_full.png"))
```

**Optimal Random Forest (Bathrooms, Rooms, Bedrooms, Stories, Kitchens, Fireplaces):**

```{r rf-optimal, cache=TRUE}
rf_optimal <- fit_optimal_rf(train, cfg)
save_importance_plot(rf_optimal, cfg, "rf_importance_optimal")
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "rf_importance_optimal.png"))
rf_eval    <- eval_rf(rf_optimal, test)
cat(sprintf("Optimal RF Test MSE: %.4f\n", rf_eval$mse))
```

---

## 4. Classification Models

### 4.1 K-Nearest Neighbors (KNN)

Classification target: **class_mean** (Reasonable / Overpriced at threshold 57.725)

```{r knn, cache=TRUE}
knn_results <- knn_loop(train_c, test_c, cfg = cfg, class_col = "class_mean")
plot_knn_error(knn_results, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "knn_error_vs_k.png"))
```

```{r knn-best}
best_k   <- knn_results$k[which.min(knn_results$mse)]
knn_best <- run_knn(train_c, test_c, k = best_k, class_col = "class_mean")
cat(sprintf("Best k = %d | Misclassification rate = %.7f\n", best_k, knn_best$mse))
```

### 4.2 Linear Discriminant Analysis (LDA)

```{r lda}
lda_model <- fit_lda(train_c, "class_mean", priors = cfg$models$lda$priors_equal)
lda_eval  <- eval_lda(lda_model, test_c, "class_mean")
cat(sprintf("LDA Correct Classification Rate: %.7f\n", lda_eval$correct_rate))
print(lda_eval$confusion)
```

### 4.3 Quadratic Discriminant Analysis (QDA)

```{r qda}
qda_model <- fit_qda(train_c, "class_mean", priors = cfg$models$qda$priors_equal)
qda_eval  <- eval_qda(qda_model, test_c, "class_mean")
cat(sprintf("QDA Correct Classification Rate: %.7f\n", qda_eval$correct_rate))
print(qda_eval$confusion)
```

```{r lda-qda-compare}
lda_qda_table <- compare_lda_qda(train_c, test_c, "class_mean", cfg)
knitr::kable(lda_qda_table, digits = 7, caption = "LDA vs QDA — Correct Classification Rates")
```

### 4.4 Support Vector Machine (SVM)

Classification target: **class_median** (Under / Over at threshold 44.365)

```{r svm-scatter}
save_svm_scatter(test_c, "class_median", cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "svm_scatter.png"))
```

```{r svm-kernels, cache=TRUE}
svm_table <- compare_svm_kernels(train_c, test_c, "class_median", cfg)
knitr::kable(svm_table, digits = 7, caption = "SVM — Kernel Comparison")
```

---

## 5. Model Comparison

```{r reg-compare}
lm_pred   <- predict(lm_model,   newdata = test)
poly_pred <- predict(poly_model, newdata = test)

reg_results <- list(
  "Linear Regression"       = list(actual = test$Price_10K, pred = lm_pred),
  "Polynomial Regression"   = list(actual = test$Price_10K, pred = poly_pred),
  "PCR"                     = list(actual = test$Price_10K, pred = pcr_eval$pred),
  "PLS"                     = list(actual = test$Price_10K, pred = pls_eval$pred),
  "Regression Tree (Pruned)"= list(actual = test$Price_10K, pred = tree_eval$pred),
  "Random Forest (Optimal)" = list(actual = test$Price_10K, pred = rf_eval$pred)
)

reg_table <- regression_comparison_table(reg_results)
knitr::kable(reg_table, digits = 4, caption = "Regression Model Comparison — Test Metrics")
```

```{r mse-bar}
save_mse_bar_chart(reg_table, cfg)
knitr::include_graphics(file.path(here(cfg$paths$plot_dir), "model_mse_comparison.png"))
```

```{r cls-compare}
svm_best_kernel <- svm_table$Kernel[which.min(svm_table$Misclassification_Rate)]
svm_best        <- fit_svm(train_c, kernel = svm_best_kernel, class_col = "class_median", cfg = cfg)
svm_eval        <- eval_svm(svm_best, test_c, "class_median")

cls_results <- list(
  "KNN (Best k)"        = list(actual = test_c$class_mean,   pred = knn_best$pred),
  "LDA (equal priors)"  = list(actual = test_c$class_mean,   pred = lda_eval$pred),
  "QDA (equal priors)"  = list(actual = test_c$class_mean,   pred = qda_eval$pred),
  "SVM (Best Kernel)"   = list(actual = test_c$class_median, pred = svm_eval$pred)
)

cls_table <- classification_comparison_table(cls_results)
knitr::kable(cls_table, digits = 7, caption = "Classification Model Comparison")
```

---

## 6. Conclusions

From the paper's findings (reproduced by this refactored pipeline):

1. **Best regression predictors**: Longitude, Bathrooms, Fireplaces, AYB_age, EYB_age, Grade
2. **Best regression model**: Polynomial regression (Adj. R² ≈ 69%, RSE ≈ 31.12)
3. **Classification models** outperform linear/logistic baselines (correct rates ~0.81)
4. **LDA** (0.8093) slightly outperforms QDA (0.8061) without priors
5. **Random Forest** most important variable: Bathrooms (optimal); Longitude (full forest)
6. **Next steps**: Combine Bayesian hierarchical models with ML ensemble methods
