---
title: "DC Property ML — Model Comparison Report"
author: "Aaron Niecestro"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    theme: readable
    toc: true
    toc_depth: 3
    number_sections: true
---

# DO NOT KNIT DOCUMENT - Works as Rmarkdown ONLY

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(yaml)
library(rprojroot)

# Find project root
root <- rprojroot::find_root(rprojroot::has_dir("config"))

# Force knitr to use project root for all file paths
knitr::opts_knit$set(root.dir = root)

# Load config
config <- yaml::read_yaml(file.path(root, "config", "config.yaml"))
cfg    <- config

# Remove trailing slash from plot_dir
plot_dir <- sub("/$", "", cfg$paths$plot_dir)

# Load data
train <- readRDS(file.path(root, cfg$paths$train_data))
test  <- readRDS(file.path(root, cfg$paths$test_data))

# Load regression comparison table
reg_table <- read.csv(file.path(root, "outputs", "tables", "run_log.csv"))

# Load classification tables
cls_table_mean   <- readRDS(file.path(root, cfg$paths$model_dir, "classification_mean_table.rds"))
cls_table_median <- readRDS(file.path(root, cfg$paths$model_dir, "classification_median_table.rds"))
```

# Executive Summary

This report evaluates machine learning models for predicting and classifying residential property prices in Washington, DC.

The analysis now includes **two separate binary classification frameworks**:

- **Mean Threshold Classification**  
  *Reasonable vs Overpriced*  
  Models: **KNN, LDA, QDA, SVM**

- **Median Threshold Classification**  
  *Underpriced vs Overpriced*  
  Models: **KNN, LDA, QDA, SVM**

Regression models continue to provide strong continuous predictions, while binary classification offers interpretable threshold‑based decision support.

---

# Regression Model Comparison

```{r}
knitr::kable(reg_table, caption = "Regression Model Comparison (Test Set)")
```

---

# Binary Classification — Mean Threshold

```{r}
knitr::kable(cls_table_mean, caption = "Mean Threshold Classification — Model Comparison")
```

### Accuracy Bar Chart

```{r echo=FALSE}
knitr::include_graphics(
  file.path(root, plot_dir, "classification_mean.png")
)
```

---

# Binary Classification — Median Threshold

```{r}
knitr::kable(cls_table_median, caption = "Median Threshold Classification — Model Comparison")
```

### Accuracy Bar Chart

```{r echo=FALSE}
knitr::include_graphics(
  file.path(root, plot_dir, "classification_median.png")
)
```

---

# Combined Comparison Table

```{r}
combined_cls_table <- dplyr::bind_rows(
  cls_table_mean %>% mutate(Threshold = "Mean"),
  cls_table_median %>% mutate(Threshold = "Median")
) %>%
  dplyr::select(Threshold, everything())

knitr::kable(combined_cls_table,
             caption = "Combined Binary Classification Comparison (Mean vs Median)")
```

---

# Side-by-Side Mean vs Median Visualization

```{r}
combined_plot_data <- dplyr::bind_rows(
  cls_table_mean %>% mutate(Threshold = "Mean"),
  cls_table_median %>% mutate(Threshold = "Median")
)

combined_plot <- ggplot(combined_plot_data,
                        aes(x = Model, y = Accuracy, fill = Threshold)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Classification Accuracy: Mean vs Median Threshold",
       x = "Model",
       y = "Accuracy") +
  theme_minimal(base_size = 14) +
  scale_fill_brewer(palette = "Set2")

ggsave(
  filename = file.path(root, plot_dir, "classification_mean_vs_median.png"),
  plot = combined_plot,
  width = 10,
  height = 6
)

knitr::include_graphics(
  file.path(root, plot_dir, "classification_mean_vs_median.png")
)
```

---

# Model Ranking Summary

```{r}
ranking_mean <- cls_table_mean %>%
  arrange(desc(Accuracy)) %>%
  mutate(Threshold = "Mean")

ranking_median <- cls_table_median %>%
  arrange(desc(Accuracy)) %>%
  mutate(Threshold = "Median")

ranking_combined <- bind_rows(ranking_mean, ranking_median) %>%
  select(Threshold, Model, Accuracy, Misclassification_Rate)

knitr::kable(
  ranking_combined,
  caption = "Model Ranking Summary — Mean vs Median Thresholds"
)
```

---

# Visual Comparison

```{r echo=FALSE}
knitr::include_graphics(
  file.path(root, plot_dir, "classification_mean.png")
)
knitr::include_graphics(
  file.path(root, plot_dir, "classification_median.png")
)
```

---

# Final Interpretation

Across all modeling approaches, ensemble tree‑based regression methods provide the most accurate continuous price predictions.  
For binary classification, SVM consistently performs well under both mean and median thresholds, while KNN benefits from neighborhood‑level structure.  
LDA and QDA offer interpretable boundaries and competitive performance under balanced class splits.  
Both threshold‑based classification frameworks provide useful segmentation depending on analytical objectives.
